upwell_IDX

### 2019-08-14 upwell_IDX.Rmd
fix merge

```{r}
south_africa_away_wide <- transects(site_list_sub, spread = 30)

world_map <- ggplot() + 
  borders(fill = "grey40", colour = "black")

# Create titles
titles <- c("Shore-normal", "Islands")

# Plotting function
plot_sites <- function(site_list, buffer, title_choice, dist){

@@

  # Visualise
  world_map +
    geom_segment(data = site_list, colour = "red4", 
                 aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
    geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
    geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
    coord_cartesian(xlim = c(min(site_list$lon - buffer), 
                             max(site_list$lon + buffer)),
                    ylim = c(min(site_list$lat - buffer), 
                             max(site_list$lat + buffer))) +
    labs(x = "", y = "", colour = "Site\norder") +
    ggtitle(titles[title_choice])
}

south_africa_away <- plot_sites(south_africa_away_wide, 1, 1, 100000)
```

```{r}
UI_angle <- wind_daily %>% 
  dplyr::rename(site = sites) %>% 
  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>% 
  dplyr::rename(coast_angle = heading) %>% 
  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # Calculated perpendicular angle for each of the sites. Not necessary
  drop_na

# index <- wind_UI %>%
#   group_by(site) %>%
#   select("date", "ui.saws") %>%
#   mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))

wind_UI <- wind_daily %>% 
  mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # 160 is angle perpendicular to the west coast of SA (See paper Fielding and Davis pg 182)
  drop_na

# Results obtained from UI_angle (Using the perpendicular angle for each site) and the results of wind_UI at -160 (angle of the west coast) appears to be the same
```

@@

```{r}
# Loading the insitu temperature data along the wc
load("Data/SACTN_US.RData")

# exceedance <-
#   function(data,
#            x = t,
#            y = temp,
#            threshold,
#            below = FALSE,
#            minDuration = 5,
#            joinAcrossGaps = TRUE,
#            maxGap = 2,
#            maxPadLength = 3)

@@

SACTN_upwell <- UI_angle %>% 
  dplyr::rename(t = date,
                temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t), 
  #        max_t = max(t)) %>% 
  nest() %>% 
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("2000-01-01", "2005-12-31")),
         exceed = purrr::map(clim, exceedance, minDuration = 3, threshold = 3)) %>% 
  # mutate() %>% 
  select(-data, -clim) %>% 
  unnest()

@@

@@

```{r}
# match_func <- function(df){
#   match <- SACTN_MCS_clims  %>% 
#   left_join(df, by = c("site", "t")) %>% 
#   na.trim()
#   return(match)
# }
# 
# index <- index %>% 
#   dplyr::rename(site = sites) %>% 
#   dplyr::rename(t = date)
# 
# match <- match_func(df = index)
# 
# upwell_MCS <- ggplot(data = match, aes(x = t, y = ui.saws)) +
#   geom_line() +
#   geom_point(data = filter(match, event == TRUE), colour = "navy") +
#   geom_rug(data = filter(match, saws.condition == "upwelling"), sides = "b") +
#   scale_x_date(expand = c(0, 0)) +
#   facet_wrap(~site, ncol = 1) +
#   theme_grey() +
#   labs(y = "Temperature (°C)", x = NULL)
# upwell_MCS

>>>>

```{r}
south_africa_away_wide <- coastR::transects(site_list_sub, spread = 30)
world_map <- ggplot() + 
  borders(fill = "grey40", colour = "black")
# Create titles
titles <- c("Shore-normal", "Islands")
# Plotting function
plot_sites <- function(site_list, buffer, title_choice, dist){

@@

  # Visualise
  world_map +
    geom_segment(data = site_list, colour = "red4", 
                 aes(x = lon, y = lat, xend = lon_dest, yend = lat_dest)) +
    geom_point(data = site_list, size = 3, colour = "black", aes(x = lon, y = lat)) +
    geom_point(data = site_list, size = 3, colour = "red", aes(x = lon_dest, y = lat_dest)) +
    coord_cartesian(xlim = c(min(site_list$lon - buffer), 
                             max(site_list$lon + buffer)),
                    ylim = c(min(site_list$lat - buffer), 
                             max(site_list$lat + buffer))) +
    labs(x = "", y = "", colour = "Site\norder") +
    ggtitle(titles[title_choice])
}
south_africa_away <- plot_sites(south_africa_away_wide, 1, 1, 100000)
```

```{r}
UI_angle <- wind_daily %>% 
  dplyr::rename(site = sites) %>% 
  left_join(south_africa_away_wide[,c(2,21)], by = "site") %>% 
  dplyr::rename(coast_angle = heading) %>% 
  mutate(ui.saws = mean_speed * (cos(dir_circ - coast_angle))) %>% # Calculated perpendicular angle for each of the sites. Not necessary
  drop_na %>% 
  dplyr::rename(t = date)
# index <- wind_UI %>%
#   group_by(site) %>%
#   select("date", "ui.saws") %>%
#   mutate(saws.condition = ifelse(ui.saws < 12.5, "upwelling"))
wind_UI <- wind_daily %>% 
  mutate(ui.saws = mean_speed * (cos(dir_circ - 160))) %>% # 160 is angle perpendicular to the west coast of SA (See paper Fielding and Davis pg 182)
  drop_na
# Results obtained from UI_angle (Using the perpendicular angle for each site) and the results of wind_UI at -160 (angle of the west coast) appears to be the same
```

@@

```{r}
# Loading the insitu temperature data along the wc
load("Data/SACTN_US.RData")
# exceedance <-
#   function(data,
#            x = t,
#            y = temp,
#            threshold,
#            below = FALSE,
#            minDuration = 5,
#            joinAcrossGaps = TRUE,
#            maxGap = 2,
#            maxPadLength = 3)

@@

SACTN_upwell <- UI_angle %>% 
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t), 
  #        max_t = max(t)) %>% 
  nest() %>% 
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("2000-01-01", "2005-12-31")),
         # NB: A threshold of 3 appeared to be far to strict
         exceed = purrr::map(clim, exceedance, minDuration = 3, threshold = 1)) %>% 
  # mutate() %>% 
  select(-data, -clim) %>% 
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% 
  dplyr::rename(ui.saws = temp) %>% 
  select(site, t, ui.saws, exceedance)

@@

# There are a few ideas on how to go about doing this
# The first is to calculatea rolling climatology and to check the upwelling phenology agianst that
# in order to determine changes in upwelling values, but this won't give us the metrics we want
# Rather what we can do is set a static bottom threshold below which we are interested in looking for upwelling
# We then use the upwelling index values created above as a second filter to determine the metrics of the upwelling
# Unfortunately the way that the heatwaveR functions work well does with purrr. 
# I've figured this out once before but can't remember where now.
# Rather than go through the process I have rather just made a convenience function below that
# does what I want and can simply be given to purrr::map() without any faffing about.
detect_event_custom <- function(df){
  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)
  return(res)
}
# Calculate the upwelling event metrics
SACTN_upwell_base <- SACTN_US %>% 
  dplyr::rename(t = date) %>% 
  group_by(site) %>% 
  nest() %>% 
  # The climatology base period used here is up for debate...
  # The choice of the 30th percentile threshold also needs to be justified and sensitivty tested
  mutate(clim = purrr::map(data, ts2clm, pctile = 30, climatologyPeriod = c("1995-01-01", "2004-12-31"))) %>%
  select(-data) %>% 
  unnest() %>%
  left_join(SACTN_upwell, by = c("site", "t")) %>%
  filter(!is.na(exceedance)) %>%
  group_by(site) %>% 
  # mutate(thresh = mean(seas, na.rm = T)-sd(seas, na.rm = T)) %>% # Manually set threshold to the mean
  # mutate(thresh = 10) %>% # Manually set threshold to a static value. Doesn't work across all sites
  # mutate(thresh = quantile(temp, 0.3, na.rm = T)) %>% # Manually set threshold to a single quantile
  nest() %>% 
  mutate(exceed = purrr::map(data, detect_event_custom)) %>% 
  select(-data) #%>% 
  # unnest() %>% 
  # filter(row_number() %% 2 == 0) %>% # Select event summary metrics
  # filter(row_number() %% 2 == 1) %>% # Select daily values
  # unnest()
# Unpack the event metric reults
SACTN_upwell_metrics <- SACTN_upwell_base %>% 
  unnest() %>%
  filter(row_number() %% 2 == 0) %>%
  unnest()
# Unpack the daily climatology results
SACTN_upwell_clims <- SACTN_upwell_base %>% 
  unnest() %>%
  filter(row_number() %% 2 == 1) %>% 
  unnest()
# The above chunk of code appears to work as expected.
# What needs to be done now is that the thresholds for duration and UI strength need to be justified/decided uopn.
# What also needs to be decided is if we are interested in any upwelling results throughout the year,
# or only during the upwelling season, whenever that may be for each site.
# I think we have demonstrated the technical capability needed to answer the question of whether or not upwelling
# is changing over time.
# Now we really need to focus on which of these parameters need to best be tweaked to answer that question.

@@

```{r}
# match_func <- function(df){
#   match <- SACTN_MCS_clims  %>% 
#   left_join(df, by = c("site", "t")) %>% 
#   na.trim()
#   return(match)
# }
# 
# index <- index %>% 
#   dplyr::rename(site = sites) %>% 
#   dplyr::rename(t = date)
# 
# match <- match_func(df = index)
# 
# upwell_MCS <- ggplot(data = match, aes(x = t, y = ui.saws)) +
#   geom_line() +
#   geom_point(data = filter(match, event == TRUE), colour = "navy") +
#   geom_rug(data = filter(match, saws.condition == "upwelling"), sides = "b") +
#   scale_x_date(expand = c(0, 0)) +
#   facet_wrap(~site, ncol = 1) +
#   theme_grey() +
#   labs(y = "Temperature (°C)", x = NULL)
# upwell_MCS
# Plot some of the Upwelling/MCS results
SP_upwelling <- SACTN_upwell_clims %>% 
  filter(site == "Sea Point", lubridate::year(t) %in% 2016:2018)
ggplot(SP_upwelling, aes(x = t, y = temp)) +
  geom_flame(aes(y = thresh, y2 = temp), fill = "navy", n = 3) +
  geom_point(data = filter(SP_upwelling, !is.na(event_no))) +
  geom_line(aes(y = thresh)) +
  geom_line(alpha = 0.7) +
  geom_rug(data = filter(SP_upwelling, event_no > 0), sides = "b")

# We can see in the figure above that the rug plot, which denotes when the wind driven upwelling was detected,
# occurrs much less frequently than simply using the 30th percentile threshold.
# The dots on the plot help to show the length of each wind driven upwelling event detected.
# This supports the use of this methodology.
# Up next is fitting linear models to the upwelling metric results to see what trends shake out.
```
